---
title: 'Data imputation'
author: "ATG"
date: "`r as.character(Sys.Date(), format = '%A %B %d, %Y')`"
output:
   rmdformats::readthedown:
      lightbox: true
      use_bookdown: true
---


```{r knitr-setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment="#>")
```

```{r output, echo = FALSE}
library(tidyverse)
```

# Introduction

* Data imputation performed using `missForest` package

# Original data

```{r assign-Rdata-nm, echo = FALSE}
data_version <- "040323"   # Date 
Rdata_nm <- paste0("olink_analytical_dataset", data_version)
```

Dataset `r Rdata_nm` stored in external file is loaded

```{r load-data}
Rdata_nm    # Original data 
fpath <- paste0("./data/", Rdata_nm,".Rdata")
load(fpath, verbose = TRUE)
```

Selected variables described for reference.

* `PRIMARY`: Primary outcome (ESKD/Dialysis, 0=NO, 1=YES)
* `YRS_PRIMARY`: Time to primary outcome or censoring (years)
* `CASE40_JUNE`: Secondary outcome (ESKD/Dialysis, or >=2 eGFR during FU or last eGFR during follow-up showing 40% decline from F04, 0=NO,1=YES )
* `YRS_CASE40_JUN`: Time to secondary outcome or censoring (years)
* `DECLINE40`: >=2 eGFR during FU showing 40% decline from F04 (0=NO, 1=YES )
* `LASTGFR_40_4`: Last eGFR during FU showing 40% decline from F04 (0=NO, 1=YES )
* `DECLINE40_PLUS`: >=2 eGFR or last eGFR during FU showing 40% decline from F04 (0=NO, 1=YES )
* `YRS_DECLINE40_PLUS`: Time to >=2 eGFR or last eGFR during FU showing 40% decline from F04  or censoring (years)
* `TM_ACCORDION`: Death from any cause (0=NO, 1=YES )
* `FU_TM_ACCORDION`: Time to death from any cause (years)
* `STATUS_PRI`:	Status for primary outcome (0=censored, 1=primary outcome, 2=death before primary outcome)
* `STATUS_SEC`:	Status for secondary outcome (0= censored, 1=secondary outcome, 2=death before secondary outcome)

Data stored in `orig_data` data frame.

```{r save-in-orig_data}
assign("orig_data", eval(as.name(Rdata_nm)))
dim(orig_data) # No of rows and vars
```

Drop rows with invalid data.

```{r drop-time0}
orig_datax <- orig_data %>% filter(YRS_PRIMARY > 0) %>% filter(YRS_CASE40_JUN > 0)
dim(orig_datax)
orig_datax$DECLINE40 <- as.factor(orig_datax$DECLINE40)
orig_datax$recno <- 1:nrow(orig_datax)
```

Show (invalid) data rows with time <= 0

```{r show-drop-data}
range(orig_data$YRS_PRIMARY)
range(orig_data$YRS_CASE40_JUN)
dt_drop <- orig_data %>% select (MASKID, SUBCO15, YRS_PRIMARY,YRS_CASE40_JUN, STATUS_PRI, STATUS_SEC) %>%
  mutate (time_0 = ifelse(YRS_PRIMARY <= 0 | YRS_CASE40_JUN <=  0, 1, 0)) %>% filter(time_0 == 1)
dt_drop
```

Glimpse `orig_datax`

```{r glimpse-orig_datax}
glimpse(orig_datax)
```


# Missing values

```{r miss-values}
miss_vals <- sapply(orig_datax, function(x) sum(is.na(x)))
miss_vals[miss_vals > 0]
names(miss_vals[miss_vals == 0]) # Variables wout missing values
```

# Var groups


* `clin_vars`: Demographic/Clinical
* `protein_vars`: Protein variables BM1, ..., BM21

```{r xvars}
all_vars <- names(orig_datax)
clin_vars <- c("AGE", "FEMALE", "BMI", "HBA1C", "BASE_GFR", "LOG_BASE_UACR","GFR_F04" )
protein_vars <- paste0("BM", 1:21)
xvars <- c(clin_vars, protein_vars)
(lenx <- length(xvars)) # No of candidate predictors
xvars # List of candidate biomarkers
tvars <- c("YRS_PRIMARY", "PRIMARY", "CASE40_JUNE","YRS_CASE40_JUN", "SUBCO15", "DECLINE40")

```

Create list of variables used by missForest for imputation

```{r missForest-variables}
(miss_vars <- c(xvars, tvars))
bmq_vars <- paste0("BMQ_", 1:21)
drvd_vars <- c(bmq_vars, "BASE_UACR") # Variables that can be derived from others
remaining_vars <- setdiff(all_vars, c(miss_vars, drvd_vars))

length(miss_vars)
length(drvd_vars)
length(remaining_vars)
length(all_vars)
```

List of vars not used by missForest

```{r remaining-vars}
drvd_vars       # Derived vars  
remaining_vars
```

# Auxiliary data


```{r tbl-miss}
orig_dtx_miss <- orig_datax %>% select(all_of(miss_vars))
orig_dtx_remaining <- orig_datax %>% select(all_of(remaining_vars))
```

```{r glimpse}
glimpse(orig_dtx_miss)
```
# MissForest for subcohort

```{r miss1-create}
library(missForest)
orig_miss1     <- orig_dtx_miss[orig_dtx_miss$SUBCO15==1, ]
orig_remaining1 <- orig_dtx_remaining[orig_dtx_miss$SUBCO15==1,]
```

```{r miss1-impute}
orig_impt1 <- missForest(orig_miss1, verbose = TRUE)
names(orig_impt1)
class(orig_impt1)
```

```{r cbind1}
dt1 <-cbind(orig_impt1$ximp, orig_remaining1) 
```

# MissForest outside of subcohort

```{r miss0-create}
#library(missForest)
orig_miss0      <- orig_dtx_miss[orig_dtx_miss$SUBCO15==0, ]
orig_remaining0 <- orig_dtx_remaining[orig_dtx_miss$SUBCO15==0,]
```

```{r miss0-impute}
orig_impt0 <- missForest(orig_miss0, verbose = TRUE)
names(orig_impt0)
class(orig_impt0)
```

```{r cbind0}
dt0 <- cbind(orig_impt0$ximp, orig_remaining0) 
```

# Combine imputed data

```{r rbind}
dtx <- rbind(dt1, dt0)
dtx_imputed <- dtx %>% arrange(recno)
dim(dtx_imputed)
names(dtx_imputed)
```

```{r Rdata_nm2}
Rdata_nm2 <- paste0("olink_analytical_imputed", data_version)
Rdpath <- paste0("./saved/", Rdata_nm2, ".Rdata")
save(dtx_imputed, file = Rdpath)
```

Create dbf file
```{r dbf}
library(foreign)
dbf_path <-  paste0("./saved/", Rdata_nm2, ".dbf")
write.dbf(dtx_imputed, dbf_path)
```

## Check missing values

```{r miss-values-1}
miss_vals <- sapply(dtx, function(x) sum(is.na(x)))
miss_vals[miss_vals > 0]
names(miss_vals[miss_vals == 0]) # Variables wout missing values
```

## Check selected variables

```{r compare}
t1 <- dtx_imputed$FEMALE - orig_datax$FEMALE
range(t1)
```
```{r knitr-knit_exit2, echo = FALSE}
knitr::knit_exit()
```


## Check var ranges


```{r Check var ranges}
orig_datax  %>% filter(if_any(where(is.numeric), ~ .x < 0)) %>%
 select(MASKID, BM6, BM19)
```


